---
title: "Evaluating Average Grocery Discounts by Store and Month in Canada in 2024"
author: 
  - Fangning Zhang
  - Cher Ning-Li
  - Kevin Roe
thanks: "Code and data are available at: [https://github.com/Kanghyunroe/grocery_discount](https://github.com/Kanghyunroe/grocery_discount)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
```


# Introduction

Since the end of the pandemic with rising inflation, grocery prices have been a contentious topic of concern for Canadians. While grocery chains have raised prices, they also have introduced various discounts to encourage shopping. 

[insert results and scope of the study]

Understanding the results of the paper suggests when what month and grocery chains has the highest average discount for each month, allowing us to make better shopping decisions. 

The paper is structured as follows: @sec-data focuses on the data and measurement methods; @sec-results focuses on the results; @sec-discussion evaluates correlation vs causation, missing data and different sources of bias. 



# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement
	
Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.



# Results {#sec-results}




# Discussion {#sec-discussion}

September highest discount

Walmart and Galleria in September

If we split each vendor have highest highest vendor - likely because of back2skool sales 

## Correlation vs Causation {#sec-discussion-correlation-and-causation}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Missing Data {#sec-discussion-missing-data}

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Sources of Bias {#sec-discussion-bias} 

## Weaknesses and next steps

Problemsiwth february becaues it starts at 28th 
\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows... 

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```



\newpage


# References


